Correspondence Analysis of Two Mode Networks

### 2025-01-29 Reviews and Response/response-memo.tex

\textbf{Response}:

\underline{Add distinction between Bonacich and CA to the highlights}:

\textbf{Reviewer}: \textit{(For example, I would like to see a ``Highlights" bullet point that says that Bonacich centrality is useful for revealing the position of the nodes in a center-periphery pattern, to the extent one exists, whereas CA is useful for revealing the presence of different groups or communities in the data, to the extent that the dual network is not homogeneous.)}

\textbf{Response}: Thanks for this suggestion. I have now included a relevant highlight bulletpoint noting the relevant analytic contrast of Bonacich versus CA.

@@

\underline{Data errors in the Southern Women two-mode matrix}:

\textbf{Reviewer}: \textit{Importantly, there are 4 errors in the matrix (``Southern Women data") that you analyze. For this dataset, you (in the note to Table 1) cite Doreian et al (2004, Table 4). Among the attendees at Event E4, you show Frances, Eleanor, and Ruth. However they did not attend this event. You also show Nora as absent from Event E8; however, Nora attended E8. These four errors are discrepant in comparison to your cited source (Doreian et al, 2004, Table 4), and they are discrepant as well in comparison to the original source (Davis, Gardener, and Gardener, 1941, p. 141, which has events in the same order as in your Table 1, and a slightly different order of women; this table is reproduced in Freeman, 2003, p. 42).}

\begin{figure}[h!]
    \includegraphics[width=1.0\textwidth]{Reviews and Response/sw-original.png}
\end{figure}

\textbf{Response}: Thanks for calling attention to this---rather embarrassing---error; it's definitely a cautionary tale against manually copying data. I have corrected the data, and as you can see in the analysis file included as supplementary material, I now extract the Southern Women data directly from the \texttt{networkdata} package. My table now matches those in the references you mentioned exactly. Note, however, that while you are correct that Frances, Eleanor, and Ruth did not attend E4, Nora did not attend E8, at least according to the published tables in Doreian et al (2004), Kovacs (2010), and the Figure 1 from Freeman (2003), shown above; in these papers, as you can see, Nora's event set is \{E6, E7, E9, E10, E11, E12, E13, E14\}. I think the confusion comes from the fact that Freeman used a non-chronological order of events (as is evident from the date labels in the figure) that emphasizes the group structure of the women. This non-chronological ordering leads to a different labeling of events than when events are arranged in chronological order; this is the ordering of events used by Doreian et al. 2004 and Kovacs (2010). In this event labeling, Nora does not attend E8 (the event that took place on 9/16.) However, note that this is obviously \textit{not} the ordering of events in the original Breiger (1974)---nor for that matter the earlier Doreian (1979) piece in \textit{Social Networks} that used Q-Analysis. In this \textit{chronological} labeling of events, Nora does attend ``E8" (the event that took place on 5/19), as evident in Breiger (1974, p. 186, fig. 2a). In any case, to forestall any confusion, and given that the in the \texttt{networkdata} package, the \texttt{igraph} object containing the Southern Women data come ordered using the Freeman (2003) ordering, I'm sticking to the non-chronological event labeling. I now include a brief discussion in the paper clarifying this issue.

### 2025-01-29 highlights.tex

\begin{itemize}
  \item This paper revisits the use of Correspondence Analysis (CA) in analyzing two-mode network data, highlighting its potential beyond data visualization.
  \item CA can compute a dual centrality score on both modes of a two-mode network, related to but not mathematically equivalent to the Bonacich (1991) two-mode centrality. 
  \item CA can extract latent positional information on people and groups, with similarities to recent work on generalized relational similarity" in two-mode networks. 
\end{itemize}

>>>>

\begin{itemize}
  \item This paper revisits the use of Correspondence Analysis (CA) in analyzing two-mode network data, highlighting its potential beyond data visualization.
  \item CA can compute a dual centrality score on both modes of a two-mode network, related to but not mathematically equivalent to the Bonacich (1991) two-mode centrality. 
  \item CA can extract latent positional information on people and groups, akin to recent work on ``generalized relational similarity" in two-mode networks. 
  \item The Bonacich dual centrality is useful for highlighting core-periphery patterns in the two-mode case in contrast to CA scores which are more appropriate to revealing the latent group structure 
\end{itemize}

### 2025-01-29 main-R1.tex

\subsection{Organization of the Paper} \label{subsec:org}
The rest of the paper is organized as follows. In the next section, I motivate the use of CA to produce a type of Bonacich-style dual centrality via weighted iteration across the two modes of the affiliation matrix. This leads naturally to an abbreviated way of computing the same scores via eigenvector decomposition of the (inverse degree-weighted) one-mode projection of the affiliation matrix for each set of nodes. This approach is computationally and mathematically related but \textit{not} equivalent to the eigenvalue decomposition of the one-mode projection matrices, which results in the usual Bonacich dual centrality scores. I then draw on recent work by \citet{van2021correspondence} to show the links between the ordering of nodes along the first CA dimension and Kovacs's generalized relational similarity, and the distance of nodes in the space formed by the first two dimensions and the connectivity-similarity between nodes across modes. I contrast this ordering to that provided by the Bonacich dual centrality scores, which, in contrast to the CA ordering, reveals a dual community partition and recovers the core-periphery structure of the two-mode network instead. In closing, I provide some pointers on how CA can be used as a method for structural (e.g., centrality and positional) analysis, independently of the usual emphasis on joint graphical displays and visual exploration.

@@

\begin{figure}[ht!]
    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
    \centering
    \includegraphics[width=0.65\textwidth]{Plots/southern-women.jpg}
    \caption{Southern Women data with rows and columns ordered as in the original DGG table.}
    \label{fig:sw}
\end{figure}

\subsection{Empirical Example} \label{subsec:ex}
\subsubsection{Souther Women Data}
The following analyses use the classic Southern Women data \citep{davis1941} (DGG) as the running example, a dataset that, according to \citet[p.]{freeman2003finding}, ``has become something of a touchstone for comparing analytic methods in social network analysis,'' having been analyzed ``again and again,'' and reappearing ``whenever any network analyst wants to explore the utility of some new tool for analyzing data.'' Freeman himself would go to review twenty-one such attempts, and the list has only gotten longer since his analytic ``meta-analysis'' \citep[e.g.,][]{doreian2004generalized, field2006identifying, roffilli2006identifying, wang2009exponential, borgatti2011analyzing, lerner2022dynamic, batagelj2022analysis,  lizardo2024two}. As we can see, novel insights can still be garnered from these old data.

One key source of confusion in the mini-literature that exists around this dataset concerns the order of presentation of the rows and columns. The reason for this is while the row mode (people) is ``nominal'' in both the scaling and the literal senses (being composed of women's names), the column mode (events) is more ambiguous. It \textit{could} be treated as nominal; however, the original DGG table includes event \textit{dates}, which means that it could also be treated as having a natural ordering, which could be reflected in the way we list the columns.  As \citet{freeman2003finding} notes, in the canonical original table, both the row and column modes had been re-ordered in a presentation to reveal what DGG thought was the group structure among the women; this means that the events were \textit{not} presented in chronological order from left to right across the columns. To add to the confusion, the original DGG table added arbitrary ordered numerical designators to both people and events, such that people came to refer to "W1" or "W11" and "E1" or "E13" based on the original DGG ordering, which may not have mattered for the women, but it mattered for the events because "E14" in the original DGG data is not the last chronological even among the fourteen columns. In fact, as shown in~\ref{tab:southern}, it is the event that occurred on August 3rd, and three events occurred after that (E4, E8, and E13) in the DGG numbering.

To further add to the confusion, some early analyses of the data, including Breiger's \citeyearpar{breiger1974duality} classic article and \citet{doreian1979evolution} used a \textit{different} ordering of the women (taken from \citet{homans}) from that of the original DGG table reproduced in \citet{freeman2003finding}, but with the events ordered \textit{chronologically}, which differs from the original DGG Table reproduced in \citet{freeman2003finding}. This means that the arbitrary ordinal labeling of the rows (e.g., "W1, "W2", "W3", etc.) and the columns (e.g., "E1, "E2", "E3", etc.) in these (and other) publications---see e.g., \citep[table 1]{doreian1979evolution}---does not correspond to the DGG ordering reproduced in \citet{freeman2003finding} and more recent publications \citep[e.g.,][]{borgatti2014analyzing, batagelj2022analysis}; different versions of the dataset reproduce different versions of the labeling. Of course, the specific labeling of the nodes does not really matter since we are usually interested in the structural properties of the rows and columns. Event ordering, however, matters for analyses that try to exploit the temporal properties of the data. Less substantively, but still importantly, the labeling of both the rows and columns does matter when trying to reconcile claims across publications from the narrative included in various papers, since "W1" or "E8" in one publication (e.g., \citet{freeman2003finding}) may not correspond to "W1" or "E8" in the other (e.g., \citet{doreian1979evolution}).

To forestall confusion, in what follows I use the numeric labeling corresponding to the original DGG data table as reproduced in \citet[Figure 1]{freeman2003finding} and \citet[Figure 28.1]{borgatti2014analyzing}, and shown in Table~\ref{tab:southern}, which include both the arbitrary ordered labels (e.g., "W1" or "E1") and both the woman's name and the event date in parentheses, with "Myrna" rebaptized to her given name of "Myra" to correspond to the original DGG table \citep{freeman2003finding, batagelj2022analysis}.

@@

\section{Illustrative Analysis of the Southern Women Data} \label{sec:anal}
\subsection{Row and Column Re-ordered Affiliation Matrices}
Figures~\ref{fig:ca-v-bon}(a) and~\ref{fig:ca-v-bon}(b) illustrate the key differences between the two versions of dual centrality $C^R$ and $C^B$. Each panel shows the original Southern Women affiliation matrix but with rows and columns re-ordered according to the first CA dimension in (a) and the \citet{bonacich1991simultaneous} centrality in (b). In the plot, a cell entry is colored red if it has a one in the original affiliation matrix and is white when the corresponding entry is zero. As we can see, the two row-column-reshuffled matrices have appreciably distinct structures, with the $C^R$ re-ordered affiliation matrix having a block-diagonal structure and the $C^B$ re-ordered affiliation matrix having a triangular structure. Accordingly, the $C^R$ centrality rankings reveal a dual community partition between groups of women who selectively attend two groups of events (on the top-right and bottom-left of the plot). The traditional eigenvector centrality re-ordering, on the other hand, reveals a classic core-periphery partitioning \citep{borgatti2000models}, with a group of highly active women (on the top-left) who selectively co-participate in highly attended events (on the top right-hand side of the plot) and less active women (in the bottom half) who go to less well-attended events (on the left side). The two centralities extract qualitatively different information from the two-mode network, with $C^R$ geared toward community partitioning and $C^B$ more focused on finding a core of well-connected people and groups.

@@

\section{Discussion and Concluding Remarks} \label{sec:disc}
This paper reconsiders the role of CA in the analysis of two-mode network data. We began with an accidental ``rediscovery'' of CA in the analysis of two-mode networks via a ``reflective'' approach \citep{hidalgo2009building}, showing that the method of reflections leads to an eigenvector-style solution that is equivalent to simple CA of the affiliation matrix. Working backward from this, we also clarified the linkages between CA and the more commonly used eigenvector approach for calculating dual centralities in two-mode data due to \citet{bonacich1991simultaneous}. I showed that just like the reflective centralities that converge to the CA scores of the affiliation matrix, we could also think of the Bonacich centralities as the equilibrium solution to reflective iterations through the two-mode matrix, with the main difference being that the CA reflections deal with sums of averages weighted by the degrees of nodes in each mode, while the Bonacich approach works with unweighted (but normalized) sums. This exercise clarifies the links between CA and the dual two-mode eigenvector centrality in a more coherent and systematic way, an issue that was left open and somewhat unclear in Bonacich's (1991) classic paper. Thus, one conclusion that emerges from this analysis is that CA computes a kind of centrality for two-mode networks, equivalent to Hidalgo and Hausman's ``reflective'' centrality \citep{van2021correspondence}.

But CA does more than reveal a latent dual ordering of nodes in the two-mode network. In addition to doing this, CA can be shown to reveal \textit{groupings} of nodes based on some conception of the similarity of their connections to nodes in the other mode. This is evident once we use the scores of the first CA dimension to re-order the rows and columns of the original affiliation matrix. When the CA scores are used, a clear (and now well-known) dual partition between persons and groups emerges in the classic Southern Women data. This partition is substantively distinct from that which would emerge if we use the same approach to re-ordering the original data using the Bonacich centralities, which instead uncover a latent core-periphery partition between actors and events.

@@

But what kind of similarity is the clustering based on the CA of the two-mode network sensitive to? Here we argued that CA recovers a type of generalized relational similarity (GRS). That is, actors who have similar patterns of linkage to similar events are deemed similar, while events that have similar patterns of connectivity to similar actors are also deemed similar. Using a well-known iterative method to recover such generalized similarities from two-mode networks \citep{jeh2002simrank}, we saw that the scores from the first dimension of CA end up being a fairly accurate approximation to the resulting partition from the generalized relational similarity approach. Thus, we can clarify that two-mode network CA reveals latent groupings of nodes based on a GRS criterion.

>>>>

\subsection{Organization of the Paper} \label{subsec:org}
The rest of the paper is organized as follows. In the next section, I motivate the use of CA to produce a type of Bonacich-style dual centrality via weighted iteration across the two modes of the affiliation matrix. This leads naturally to an abbreviated way of computing the same scores via eigenvector decomposition of the (inverse degree-weighted) one-mode projection of the affiliation matrix for each set of nodes. This approach is computationally and mathematically related but \textit{not} equivalent to the eigenvalue decomposition of the one-mode projection matrices, which results in the usual Bonacich dual centrality scores. I then draw on recent work by \citet{van2021correspondence} to show the links between the ordering of nodes along the first CA dimension and Kovacs's generalized relational similarity, and the distance of nodes in the space formed by the first two dimensions and the connectivity-similarity between nodes across modes. I contrast this ordering to that provided by the Bonacich dual centrality scores, which, in contrast to the CA ordering, reveals a dual community partition and recovers the core-periphery structure of the two-mode network instead \citep{everett2013dual}. In closing, I provide some pointers on how CA can be used as a method for structural (e.g., centrality and positional) analysis, independently of the usual emphasis on joint graphical displays and visual exploration.

\section{Previous uses of CA and MCA in Two-Mode Network Data Analysis}

\subsection{CA as a Data Visualization Tool}

\citet{borgatti1997network} provided an early in-depth considerations of CA in the context of two-mode network data analysis with an emphasis on the visualization aspect. The discussion was placed in a section on ``visual representations of 2-mode data,'' and naturally focused almost exclusively on the uses of CA as visualization technique. According to \citet[246]{borgatti1997network}, CA ``can be seen as a method for representing points in a metric space such that distances between the points are meaningful.'' More specifically, \citet{borgatti1997network}, also using the Southern Women data as illustration,  claimed that in the CA map, points representing the women will appear closer if the women attend ``mostly the same events,'' and conversely points representing the events will be placed near one another if the events were attended by ``mostly the same women,'' and that women and events would appear closer in the plot if ``those women attended those events.'' \citet[247-249]{borgatti1997network} then went on to focus on various limitations of CA as visualization tool, noting that given the limiting range of dichotomous network data, two-dimensional representations ``will almost always be severely inaccurate or misleading.'' They also noted that since CA is a technique designed to deal with valued frequency data, its use with binary network data would lead to dimensional scores that would be either ``meaningless'' or ``difficult to interpret.'' Finally, they noted that since distances in the CA biplot are not Euclidean they would mislead interpreters who would try to impose such a metric to make sense of them.

In a paper published shortly thereafter \citet{roberts2000correspondence}, responded to some of the criticisms that \citet{borgatti1997network} levied at the use of CA for analyzing two-mode networks in their original piece. First, \citet{roberts2000correspondence} noted that there were many examples in the standing of using CA to analyze binary non-frequency data, and that therefore the uses of CA are not constrained by the nature or provenance of the numerical quantities recorded in each cell of a two-mode network affiliation matrix. \citet{roberts2000correspondence} also challenged the contention that the two-dimensional representation of the data was unduly limiting in its capacity to produce a faithful representation of the original data, noting a high correlation between the distances between the women and and the event points in the two-dimensional representation and one that used all thirteen dimensions. Finally, \citet{roberts2000correspondence} also noted that the Euclidean distance interpretation of the separation of women and events points in the joint display is partially appropriate, as this Euclidean distances as actually low-dimensional approximations (to a high degree of accuracy in the first few dimensions) to the $\mathcal{X}^2$ distances between row profiles (for the women) and column profiles (for events) where the row profiles for a given woman is equal to the entry in the original affiliation matrix divided by the square root of the product of the number of events attended by that woman and the number of women who attended that even (and similarly for column profiles corresponding to events). Thus, the point distances are indeed interpretable as a kind of degree-weighted similarity between objects in each mode.

\subsection{CA and Two-Mode Centrality}
One of the earliest considerations of CA as a data-analytic tool for two-mode networks was provided by \citet[163-165]{bonacich1991simultaneous} in his classic paper on dual centralities for affiliation networks. In that paper \citet{bonacich1991simultaneous} noted that since the usual affiliation matrix is a ``cross-classification of people buy the groups to which they belong,'' then it would seem that CA would be an appropriate analytic technique. According to \citet{bonacich1991simultaneous}, the basic goal of correspondence analysis is to create a multidimensional space encoding geometric ``distances'' (square quotes in the original) between row and column categories, with the distances capturing the best low-dimensional representation (in the least-squared error) sense of the similarities between row and column objects.  Bonacich's discussion of CA came after his influential introduction of the Eigenvector centrality approach for two-mode network analysis. Intriguingly, \citet[163]{bonacich1991simultaneous} noted that ``[a]lthough the goals of correspondence analysis and centrality analysis are different, their mathematics are similar,'' because both used a version of the spectral decomposition of a matrix---e.g., the Singular Value Decomposition (SVD).

As we will see later, the mathematics connecting the Eigenvector approach and CA are more similar than even Bonacich appreciated. Regardless, for Bonacich, the differences between Eigenvector centrality analysis and CA stemmed from ``the different ways in which the cross-classifications are treated before the SVD is applied.'' For Bonacich, the ``most powerful and useful dimensions'' of CA are designed to highlight ad describe \textit{similarities} in row or column patterns of affiliation and membership, and therefore ``[b]before the SVD is used, the cross-classification is modified so that these dimensions of similarity will be the most powerful.'' Nevertheless, Bonacich is not quite clear as to what this pre-modification amounts to, noting that given the fact that less ``powerful'' dimensions are usually discarded, it could be that among these ``there may be a dimension corresponding to centrality," but ``correspondence analysis is not designed to highlight this dimension.'' Interestingly, Bonacich went on to compute both the Eigenvector and CA scores for the canonical Southern Women data---which we will also use in this paper---noting that ``[r]rather than centrality what\ldots[the CA scores] seem to capture is membership in the two cliques that attended different sets of events.'' The magnitude of each woman and event score in the main dimension seemed to indicate the ``purity'' (or centrality!) of their membership in each of the two cliques, and for events their likelihood of being attended by the most representative members of them 
\citep[164]{bonacich1991simultaneous}.

In an influential paper on centrality in two-mode networks \citet{faust1997centrality} returned to the issue of the connection between eigenvector centrality and CA. Faust noted that one thing both techniques had in common is that they both resulted in scores that truly exploited the duality of the data. In the eigenvector case, this duality is strict, as the eigenvector centrality of a given person is the sum of the eigenvector scores of the even she attends, and for events their scores are given by the sum of the eigenvector centralities of its members. Importantly, \citet{faust1997centrality} noted that an exact parallel dual additive relationship obtained between the CA scores corresponding to persons and groups. The CA score of a given person on a given dimension is the (weighted) sum of the scores in the same dimension of the events she attends, and for groups, the score on a given CA dimension is the (weighted) sum of the scores of the women who attend that event. \citet{faust1997centrality} also specified the particular way that the affiliation matrix $\mathbf{A}$ had to be modified before applying the SVD (recall this is the step that creates the difference between CA and Eigenvector centrality), noting that each cell $a_{pg}$ is divided by the square root of the product of the number of events the row woman attends and the number of people attending the column event ($\sqrt{\sum_p a_{pg}\sum_g a_{pg}}$).

@@

\begin{figure}[ht!]
    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
    \centering
    \includegraphics[width=0.65\textwidth]{Plots/southern-women.jpg}
    \caption{Southern Women data matrix with rows and columns ordered as in the original DGG table.}
    \label{fig:sw}
\end{figure}

\subsection{Empirical Example} \label{subsec:ex}
\subsubsection{Southern Women Data}
The following analyses use the classic ``Southern Women'' data \citep{davis1941} (DGG) as the running example. This data is doubly appropriate for showcasing the relevance of methods that exploit the duality between two sets of entities, as it was it used in Breiger's \citeyearpar{breiger1974duality} classic paper. Moreover, Southern Women is dataset that, according to \citet[p.]{freeman2003finding}, ``has become something of a touchstone for comparing analytic methods in social network analysis,'' having been analyzed ``again and again,'' and reappearing ``whenever any network analyst wants to explore the utility of some new tool for analyzing data.'' Freeman himself would go to review twenty-one such attempts, and the list has only gotten much longer in the last two decades since Freeman's methodological ``meta-analysis'' \citep[e.g.,][among many others]{doreian2004generalized, field2006identifying, roffilli2006identifying, wang2009exponential, kovacs2010generalized, brusco2011analysis, borgatti2014analyzing, everett2013dual, lerner2022dynamic, batagelj2022analysis, lizardo2024two}. Despite this long list of studies of the Southern Women data, we will see that novel insights can still be garnered from this old data source.

Despite its undisputed pedigree as a methodological touchstone in studies of duality and dual constitution in two-mode networks, a key source of confusion in the mini-literature that exists around the Southern Women data concerns the order of presentation of the rows and columns. The reason for this is while the row mode (people) is ``nominal'' in both the scaling and the literal senses (being composed of women's names), the column mode (events) is more ambiguous. It \textit{could} be treated as nominal; however, the original DGG table includes event \textit{dates}, which means that it could also be treated as having a natural ordering, which could be reflected in the way we list the columns.  As \citet{freeman2003finding} notes, in the canonical original table, both the row and column modes had been re-ordered in a presentation to reveal what DGG thought was the group structure among the women; this means that the events were \textit{not} presented in chronological order from left to right across the columns.

To add to the confusion, the original DGG table added arbitrary ordered numerical designators to both people and events, such that people came to refer to "W1" or "W11" and "E1" or "E13" based on the original DGG ordering, which may not have mattered for the women, but it mattered for the events because "E14" in the original DGG data is not the last chronological even among the fourteen columns. In fact, as shown in Figure~\ref{fig:sw}, it is the event that occurred on August 3rd, and three events occurred after that (E4, E8, and E13) in the DGG numbering.\footnote{Using the DGG labeling, the chronologically ordered set of events are \{E11, E5, E2, E7, E12, E9, E3, E6, E10, E1, E14 E8, E4, E13\} \citep[p. 68]{everett2018measuring}.}

To further add to the confusion, some early analyses of the data, including Breiger's \citeyearpar{breiger1974duality} classic article and \citet{doreian1979evolution} used a \textit{different} ordering of the women---taken from \citet[83]{homans1950human}---from that of the original DGG table reproduced in \citet{freeman2003finding}, and ordered the events from left to right in the affiliation matrix \textit{chronologically}, which, as we have seen, also differs from the original DGG Table reproduced in \citet{freeman2003finding}. This means that the arbitrary ordinal labeling of the rows (e.g., "W1, "W2", "W3", etc.) and the columns (e.g., "E1, "E2", "E3", etc.) in these (and other) publications---see e.g., \citep[table 1]{doreian1979evolution}---does not correspond to the DGG ordering reproduced in \citet{freeman2003finding} and more recent publications \citep[e.g.,][]{borgatti2014analyzing, batagelj2022analysis}; in all, different versions of the dataset reproduce different versions of the labeling.

Of course, the specific labeling of the nodes does not really matter since we are usually interested in the structural properties of the rows and columns. Event ordering, however, matters for analyses that try to exploit the temporal properties of the data \citep[e.g.,][]{everett2018measuring, lerner2022dynamic}. Less substantively, but still importantly, the labeling of both the rows and columns does matter when trying to reconcile claims across publications from the narrative included in various papers, since "W1" or "E8" in one publication \citep[e.g.,][]{freeman2003finding} may not correspond to "W1" or "E8" in the other \citep[e.g.,][]{doreian1979evolution}.

To forestall confusion, in what follows, I use the numeric labeling corresponding to the original DGG data table as reproduced in (among others) \citet[Figure 1]{freeman2003finding} and \citet[Figure 28.1]{borgatti2014analyzing}, and shown in Table~\ref{fig:sw}. My presentation includes both the arbitrary ordered labels (e.g., "W1" or "E1") and both the woman's name and the event date in parentheses, with "Myrna" (W11) rebaptized to her given name of "Myra" to correspond to the original DGG table \citep{freeman2003finding, batagelj2022analysis}.

@@

\section{Illustrative Analysis of the Southern Women Data} \label{sec:anal}
\subsection{Row and Column Re-ordered Affiliation Matrices}
Figures~\ref{fig:ca-v-bon}(a) and~\ref{fig:ca-v-bon}(b) illustrate the key differences between the two versions of dual centrality $C^R$ and $C^B$. Each panel shows the original Southern Women affiliation matrix but with rows and columns re-ordered according to the first CA dimension in (a) and the \citet{bonacich1991simultaneous} centrality in (b). In the plot, a cell entry is colored red if it has a one in the original affiliation matrix and is white when the corresponding entry is zero. As we can see, the two row-column-reshuffled matrices have appreciably distinct structures, with the $C^R$ re-ordered affiliation matrix having a block-diagonal structure and the $C^B$ re-ordered affiliation matrix having a triangular structure. Accordingly, the $C^R$ reflective scores reveal a dual community partition between groups of women who selectively attend two groups of events (on the top-right and bottom-left of the plot).

The traditional eigenvector centrality re-ordering, on the other hand, reveals a classic core-periphery partitioning \citep{borgatti2000models}, with a group of highly active women (on the top-left) who selectively co-participate in highly attended events (on the top right-hand side of the plot) and less active women (in the bottom half) who go to less well-attended events (on the left side). In fact, as \citet[p. 206-207]{everett2013dual} have noted, computing the eigenvector scores for both sets of nodes in a the mode network from the classic Breiger projections---as in \citet{bonacich1991simultaneous}---and using them to re-order the rows and columns of the original affiliation matrix is a way to approximate using continuous scores an ideal discrete core-periphery partitioning of the two-mode data matrix, cheaper and more efficiently compared to direct optimization methods.  The larger point is that the standard eigenvector and reflective scores extract qualitatively different information from the two-mode network, with $C^R$ geared toward community partitioning and $C^B$ more focused on finding a core of well-connected people and groups.

@@

\section{Discussion and Concluding Remarks} \label{sec:disc}
This paper reconsiders the role of CA in the analysis of two-mode network data. We began with an accidental ``rediscovery'' of CA in the analysis of two-mode networks via a ``reflective'' approach \citep{hidalgo2009building}, showing that the method of reflections leads to an eigenvector-style solution that is equivalent to simple CA of the affiliation matrix. Working backward from this, we also clarified the linkages between CA and the more commonly used eigenvector approach for calculating dual centralities in two-mode data due to \citet{bonacich1991simultaneous}. I showed that just like the reflective centralities that converge to the CA scores of the affiliation matrix, we could also think of the Bonacich centralities as the equilibrium solution to reflective iterations through the two-mode matrix, with the main difference being that the CA reflections deal with sums of averages weighted by the degrees of nodes in each mode, while the Bonacich approach works with unweighted (but normalized) sums. This exercise clarifies the links between CA and the dual two-mode eigenvector centrality more coherently and systematically, an issue that was left open and somewhat unclear in Bonacich's \citeyearpar{bonacich1991simultaneous} classic paper. Thus, one conclusion that emerges from this analysis is that CA computes a kind of centrality for two-mode networks, equivalent to Hidalgo and Hausman's ``reflective'' centrality \citep{van2021correspondence}.

But CA does more than reveal a latent dual ordering of nodes in the two-mode network. In addition to doing this, CA can be shown to reveal \textit{groupings} of nodes based on some conception of the similarity of their connections to nodes in the other mode. This is evident once we use the scores of the first CA dimension to re-order the rows and columns of the original affiliation matrix. When the CA scores are used, a clear (and now well-known) dual partition between persons and groups emerges in the classic Southern Women data. This partition is substantively distinct from that which would emerge if we use the same approach to re-ordering the original data using the Bonacich centralities, which instead uncover a latent core-periphery partition between actors and events \citep{everett2013dual}.

@@

But what kind of similarity is the clustering based on the CA of the two-mode network sensitive to? Here, we show that CA recovers a type of generalized relational similarity (GRS). That is, actors who have similar patterns of linkage to similar events are deemed similar, while events that have similar patterns of connectivity to similar actors are also deemed similar. Using a well-known iterative method to recover such generalized similarities from two-mode networks \citep{jeh2002simrank}, we saw that the scores from the first dimension of CA end up being a fairly accurate approximation to the resulting partition from the generalized relational similarity approach. Thus, we can clarify that two-mode network CA reveals latent groupings of nodes based on a GRS criterion.
